\vspec{} and Friends: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-07-19 src/tex/ms.tex
Complete VSPEC and pypsg sections

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage[version=4]{mhchem}

@@

The flow of \vspec{}'s code is broken into three parts:
\begin{enumerate}
    \item Read in configuration \\
    \item Compute planetary spectra \\
    \item Compute stellar spectra
\end{enumerate}

The main object that a user interacts with is the \texttt{ObservationModel}. This object can be initialized by passing it either a YAML file containing
all of the simulation information, or a \texttt{InternalParameters} object encoded with the same information. Upon initialization a local working directory
(by convention \texttt{.vspec}) is created to store simulation outputs and intermediate files.

The second step is initiated by calling the \texttt{ObservationModel.build\_planet()} method. This is where all the interaction with PSG takes place.
First, your 3D Global Circulation Model (GCM) is uploaded to PSG via the API's \texttt{set} method. Then, PSG is updated with information about the observation
that does not change with time, such as instrument parameters and system information. Before entering the main loop \vspec{} computes the geometry of the system
at each epoch, storing it for later. \vspec{} then iterates through each phase, making four calls to the PSG API:
\begin{enumerate}
    \item An \texttt{upd} request setting the geometric parameters for this epoch. \\
    \item An \texttt{all} request to retrieve the combined light spectrum. \\
    \item A second \texttt{upd} request to set the stellar flux to zero. \\
    \item A second \texttt{all} request to retrieve the thermal emission of the planet.
\end{enumerate}

Results from these calls are stored locally in the \texttt{.fits} format.

The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. First, this model is initialized based on the user-passed parameters.
One consideration made during this step is that a user may want a shorter cadence here than is necessary to resolve temporal features in the planetary spectrum,
so the geometry is recalculated and the planetary spectrum at each epoch is a linear interpolation of the results from the previous step. After this setup
\vspec{} enters the main loop, where it computes the composite stellar spectrum based on the stellar model and spectral grid. For each epoch it returns this stellar
spectrum as seen by the observer, the spectrum as seen by the planet, the reflected spectrum, the planetary thermal spectrum, the total combined spectrum, and the
total noise spectrum. The data at each epoch is stored in a \texttt{.fits} file.

The final outputs of \vspec{} should be easily read using Astropy's \texttt{QTable} object, but we have also included a class \texttt{PhaseAnalyzer} to automatically
parse all output files and conveniently access 1D spectra or lightcurves.

>>>>

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage[version=4]{mhchem}
\usepackage{listings}

@@

The flow of \vspec{}'s code is broken into three parts: i) read in the configuration, ii) compute planetary spectra, and iii) compute stellar spectra.
\subsection{Configuring \vspec{}}
\label{subsec:config}

\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
be written:
\begin{lstlisting}[label={ls:yaml}]
system:
    distance: 12.4 pc
    inclination: 89.7 deg
    phase_of_pariasteron: 0 deg
\end{lstlisting}
Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.

The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
parameters in Listing \ref{ls:yaml} would be written:
\begin{lstlisting}[language=Python]
from astropy import units as u
from VSPEC.params import *
params = InternalParameters(
    ... # Other arguments
    system=SystemParameters(
        distance=12.4*u.pc,
        inclination=89.7*u.deg,
        phase_of_pariasteron=0*u.deg
    )
)
\end{lstlisting}
This is convenient for producing configurations programically, for example for producing a grid of model phase curves.

However the user decides to provide model parameters, they are read into the main \vspec{} object: the \texttt{ObservationModel}.
This is the class that runs the model and handles files. Upon initialization a local working directory
(by convention \texttt{.vspec}) is created to store simulation outputs and intermediate files.

\subsection{Planetary Phase Curve}
\label{subsec:phase-curve}
\vspec{} generates a phase curve though a series of API calls to the Planetary Spectrum Generator \citep[PSG][]{villanueva2018}, using
\texttt{pypsg} (see Section \ref{sec:pypsg}) to interface with the API. Initially, configurations are sent that give PSG information about
the system that does not change with time -- that includes the 3D Global Circulation Model (GCM) and instrument parameters. \vspec{} then enters
the main loop, iterating through each observation epoch and making a pair of API calls: one to get a combined light spectrum and one to get the thermal flux only.
Results from PSG in each epoch are stored locally in the \texttt{.fits} format.

\subsubsection{Phase Sampling}
Making API calls to PSG can be computationally expensive, especially when trying to resolve small structures in the climate model. In order to reduce unnecessary
calls to PSG the temporal sampling of the planetary spectrum in independent of the stellar model/output sampling. When computing the planetary flux in the output files
the raw PSG output is interpolated and averaged over the integration using the trapazoid rule. By default one spectrum is computed at each interface separating integrations.

\subsection{Stellar Lightcurve}

The last step is to replace the stellar spectra used by PSG with \vspec{}'s stellar model. In this step \vspec{} acts as a bridge between
the stellar surface model \texttt{vspec-vsm} (which models spatial changes to the star with time, see Section \ref{sec:star}) and a grid of
pre-computed stellar spectra (see Section \ref{sec:gridpolator}). The star is initialized and allowed to evolve for a user-specified amount of time
in order for spots and faculae to approach growth-decay equilibrium (this is also called the ``burn in'' phase). The star is then evolved at the output cadence,
at each epoch generating a set of key-value pairs that describe the surface temperature and coverage fractions of the portion of the stellar disk visible to the observer.
These \teff\ s and coverage fractions are fed into GridPolator to produce composite stellar spectra.

\subsubsection{Reflected light}
\vspec{} also produces a composite spectrum of the
portion of the star visible to the planet to accurately incorporate the variable star into the reflected light spectrum. To compute the total reflected light
in the output files \vspec{} first divides the reflected light flux from PSG by the PSG stellar model to obtain the apparent albedo. This albedo is then multiplied
by the composite spectrum (facing the planet) to obtain total reflected light flux.

\subsubsection{Noise}
Alongside the planetary radiance files PSG returns a file giving the noise profile for each epoch broken down by source. We assume that the total noise is the
quadrature sum of its constituent parts and that the \texttt{source} column (i.e. photon noise) is the only one effected by replacing one stellar spectrum with the other.
This means the photon noise can be calculated:
\begin{equation}
    N_{\rm photon} = N_{\rm photon,~PSG} \sqrt{\frac{F_{*,\rm ~VSPEC}}{F_{*,\rm ~PSG}}}
\end{equation}
where $F_*$ is the stellar flux. The total noise associated with the integration is then
\begin{equation}
    N_{\rm tot} = \sqrt{
        N_{\rm photon}^2 + N_{\rm detector}^2 + N_{\rm telescope}^2 + N_{\rm background}
    }
\end{equation}

\subsubsection{Transit and Eclipse}
Users must be careful when using \vspec{} to simulate transit and eclipse measurements because of the vast timescale differences between those events
and the typical phase variations of the planet -- a sparsely sampled phase curve that happens to include one epoch of transit will not describe that lightcurve accurately
because the interpolator knows no better than to assume a linear interpolation scheme. It is recommended that users use a high cadence for observations of transits
and eclipses so that each event will be sampled by more than one point.

In the case of a transiting geometry \vspec{} computes the spectrum of the occulted portion of the star in order to properly simulate the
transit light source effect \citep[TLS][]{rackham2018}. This spectrum is the flux blocked by the planet in the case that it is a bare rock (i.e. a flat
transmission spectrum). The effect of atmospheric transmission is added by comparing the PSG-computed effective radius to a purely geometric
calculation of a bare-rock transit depth. That is
\begin{equation}
    F_{\lambda} = F_{{\rm rock}, \lambda}\, F_{{\rm PSG,} \lambda} \,\left (\frac{R_*}{R_p} \right )^2
\end{equation}
where $F_\lambda$ is the flux of the variable star surface occulted by the planet with the atmosphere considered. $F_{{\rm rock}, \lambda}$ is the
flux of the surface occulted by a bare-rock with the planet's radius, $F_{{\rm PSG,} \lambda}$ is the flux computed by PSG to be lost
due to occultation and transmission, and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.

In the case of a total eclipse of the planet the thermal and reflected spectra are set to zero flux. However, in the case of a partial eclipse,
\vspec{} computes the fraction of the planet that is visible to the observer and reduces both thermal and reflected spectra accordingly. In this case
\vspec{} treats the planet homogeneously, however future work could utilize PSG/GlobES's ability to return a hypercube of spatially resolved
spectra in order to construct planetary spectra only from regions of the planet that are not eclipsed.

\subsection{Analysis Tools}
The final outputs of \vspec{} should be easily read using Astropy's \texttt{QTable} object, but we have also included a class \texttt{PhaseAnalyzer} to automatically
parse all output files and conveniently access 1D spectra or lightcurves. \TJ{Is it worth talking about this?}

\section{Interfacing with the Planetary Spectrum Generator via \texttt{pypsg}}
\label{sec:pypsg}

PSG\footnote{\url{https://psg.gsfc.nasa.gov/}} \citep{villanueva2018} is a powerful radiative transfer tool that is ubiquitous in the exoplanet and solar system atmosphere fields.
Either though its web graphical interface, its public API, or the local version available through Docker, PSG allows users to simulate observations of planets, comets, and
moons with a variety of geometries and realistic noise models \TJ{probably someone like Geronimo should write this part}.

During development of \vspec{} it became clear that a Python interface for the PSG API system would be generally useful to planetary scientists and astronomers, and
so \vspec{}'s own interface was spun off into a stand-alone library: \texttt{pypsg}.

\texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django to encode a PSG configuration file as a native Python object.
Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information.

Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's
\texttt{requests} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
could be committed to a public repository.

\subsection{GlobES}
\label{subsec:globes}

A powerful feature of \texttt{pypsg} is that it provides a universal interface for 3D atmosphere models and PSG's Global Emission Spectra (GlobES) application.
GlobES expects 3D data to be presented in a binary data format, and converting a model to this format is non-trivial. \texttt{pypsg}'s \texttt{PyGCM} class provides
a native Python representation of GlobES input, and includes built-in methods for converting popular climate models (e.g. ExoCAM, WACCM \TJ{full names and ref?}) to
the correct format.
